---
output:
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    slide_number: true
    transition: fade
    center: false
    css: custom.scss
    fig_width: 11
    fig_height: 6
    fig_retina: 2
---

## 

```{r title-slide, echo=FALSE, fig.width=12, fig.height=9}
#| message: false

par(mar = c(1, 1, 1, 1), bg = "#f5f5f0")
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10))

# Top decorative line
lines(c(1, 9), c(9.5, 9.5), col = "#8B0000", lwd = 2)
lines(c(1, 9), c(9.3, 9.3), col = "#8B0000", lwd = 0.8)

# Main title (largest)
text(5, 8.5, "The State of Food Security", cex = 2.2, font = 2, col = "#8B0000")
text(5, 7.8, "and Nutrition in the US", cex = 2.2, font = 2, col = "#8B0000")

# Subtitle (medium, italic)
text(5, 6.8, "A Crisis in Our Backyard", cex = 1.4, font = 3, col = "#555555")

# Decorative separator
lines(c(3.5, 6.5), c(6.2, 6.2), col = "#D3D3D3", lwd = 0.8)

# Author and date (smaller)
text(5, 5.5, "Mehreen Ali Gillani", cex = 1.2, font = 2, col = "#333333")
text(5, 5.0, "03 May 2026", cex = 1.0, col = "#666666")

# Divider
lines(c(2, 8), c(4.4, 4.4), col = "#D3D3D3", lwd = 0.5, lty = 2)

# What's included (small, informative)
text(5, 4.0, "This Analysis Examines:", cex = 0.85, font = 2, col = "#333333")
text(5, 3.5, "• 51-state poverty & food insecurity correlation (r = 0.985)", cex = 1.1, col = "#555555")
text(5, 3.1, "• Geographic 'Hunger Belt' and regional disparities", cex = 1.1, col = "#555555")
text(5, 2.7, "• Intergenerational impact on children becoming adults", cex = 1.1, col = "#555555")
text(5, 2.3, "• Policy recommendations for Congressional action", cex = 1.1, col = "#555555")

# Bottom accent
rect(3.5, 1.2, 6.5, 1.4, col = "#8B0000", border = NA)

# Footer note
text(5, 0.8, "Data: USDA Economic Research Service & Census Bureau 2023", cex = 0.9, col = "#999999")
```

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggrepel)

df <- read.csv('data/master_dataset.csv')

state_abbr <- data.frame(
  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
            "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia", 
            "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", 
            "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", 
            "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", 
            "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", 
            "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", 
            "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", 
            "Washington", "West Virginia", "Wisconsin", "Wyoming"),
  abbr = c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", 
           "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", 
           "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", 
           "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", 
           "UT", "VT", "VA", "WA", "WV", "WI", "WY")
)

df <- df %>% left_join(state_abbr, by = "state")
hunger_belt <- c("Louisiana", "Mississippi", "Arkansas", "Alabama", 
                 "Kentucky", "West Virginia", "Oklahoma")
df <- df %>% mutate(region = ifelse(state %in% hunger_belt, "Hunger Belt", "Other States"))

# Calculate all values in setup so they're available globally
cor_val <- round(cor(df$poverty_all, df$food_insecurity_proxy), 3)
nat_avg <- mean(df$food_insecurity_proxy)
```

## The Myth

#### Food Insecurity is a Crisis in the United States

```{r}
#| label: Food-Insecurity-Crisis
#| echo: false
#| eval: true
#| fig-show: hold
#| message: false
#| fig-width: 11
#| fig-height: 7
#| fig-cap: "Food Insecurity is a Crisis in the United States"
par(mfrow = c(1, 2), mar = c(2, 1, 3, 1), bg = "#f5f5f0")

plot(1, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10))
text(5, 7, "USA", cex = 3.5, font = 2)
text(5, 5, paste0(round(mean(df$food_insecurity_proxy), 1), "%"), cex = 5, col = "#8B0000", font = 2)
text(5, 3.2, "Food Insecure", cex = 1.3)
title("IN OUR BACKYARD", col.main = "#8B0000", font.main = 2, cex.main = 1.2)

plot(1, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10))
text(5, 7, "Global", cex = 3.5, font = 2)
text(5, 5, "9.8%", cex = 5, col = "#666666", font = 2)
text(5, 3.2, "Food Insecure", cex = 1.3)
title("THE 'ELSEWHERE' MYTH", col.main = "#666666", cex.main = 1.2)

mtext("US rate (8.7%) nearly equals global average (9.8%)", 
      side = 1, line = 0, cex = 1.1, font = 2, col = "#8B0000")
```      

## The Core Truth
#### Poverty and Hunger Are Inseparable


```{r Poverty-vs-Hunger-all-states, echo=FALSE, fig.width=10, fig.height=6}
#| message: false

ggplot(df, aes(x = poverty_all, y = food_insecurity_proxy)) +
  geom_point(color = "#8B0000", size = 3, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "#2c3e50", linewidth = 1.2, fill = "#f0f0f0") +
  geom_text_repel(aes(label = state),  # All states, not just filtered
                  size = 2.5,
                  alpha = 0.6,
                  max.overlaps = 20,
                  segment.size = 0.2) +
  labs(x = "Poverty Rate (%)", 
       y = "Food Insecurity Rate (%)",
       title = "Poverty = Hunger",
       subtitle = paste("r =", cor_val, "| 51-State Analysis")) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#8B0000", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 10))
```



## The Hunger Belt

#### 7 States = 42% of Food Insecure Children
```{r hunger-belt, echo=FALSE, fig.width=10, fig.height=5}
#| message: false

df_summary <- df %>%
  group_by(region) %>%
  summarise(
    avg_food_insecurity = mean(food_insecurity_proxy),
    total_children = sum(children_food_insecure_est, na.rm = TRUE)
  )

ggplot(df_summary, aes(x = region, y = avg_food_insecurity, fill = region)) +
  geom_bar(stat = "identity", width = 0.4) +  # Smaller bars
  geom_text(aes(label = paste0(round(avg_food_insecurity, 1), "%")), 
            vjust = -0.4,  # Move text higher to avoid overlap
            size = 4.5,    # Smaller font (was 8)
            fontface = "bold") +
  scale_fill_manual(values = c("#8B0000", "#A9A9A9")) +
  labs(x = "", y = "Food Insecurity Rate (%)",
       title = "Hunger Belt: 3.4x Higher Food Insecurity",
       subtitle = "LA, MS, AR, AL, KY, WV, OK",
       caption = paste("Children affected:", format(df_summary$total_children[1], big.mark = ","))) +
  theme_minimal(base_size = 10) +  # Smaller base font
  theme(legend.position = "none",
        plot.title = element_text(face = "bold", hjust = 0.5, color = "#8B0000", size = 12),
        plot.subtitle = element_text(hjust = 0.5, size = 9, face = "bold"),
        plot.caption = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 8))
```

## Hidden in Plain Sight: Millions of Children in Poverty

#### Raw numbers tell a different story than percentages

```{r Hidden-in-Plain-Sight-clean, echo=FALSE, fig.width=9, fig.height=6}
#| message: false

# Label only top 7 states by children in poverty (matches Hunger Belt narrative)
top_states <- df %>%
  arrange(desc(children_in_poverty)) %>%
  slice(1:7) %>%
  pull(state)

# Calculate total children for these 7 states in MILLIONS
total_children_7 <- df %>%
  filter(state %in% top_states) %>%
  summarise(total = sum(children_in_poverty)) %>%
  pull(total) / 1000000  # Convert to MILLIONS

ggplot(df, aes(x = population_2023 / 1e6, y = children_in_poverty / 1e3)) +
  geom_point(aes(color = poverty_children_u17), alpha = 0.85, size = 4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8B0000", linewidth = 0.8, linetype = "dashed") +
  geom_text_repel(
    data = df %>% filter(state %in% top_states),
    aes(label = state), 
    size = 3.5, 
    fontface = "bold", 
    color = "#8B0000",
    box.padding = 0.5
  ) +
  scale_color_gradientn(name = "Child Poverty Rate (%)",
                        colors = c("#2ecc71", "#f39c12", "#c0392b")) +
  scale_x_log10(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma_format(scale = 1, suffix = "K")) +
  labs(
    title = paste("7 States Contain", round(total_children_7, 1), "Million Children in Poverty"),
    subtitle = "CA, TX, FL, NY, IL, PA, OH account for 40% of all poor children",
    x = "State Population (log scale)",
    y = "Children in Poverty (Thousands)"
  ) +
  annotate("text", x = 30, y = 200, 
           label = "Target these 7 states\nfor maximum impact", 
           size = 3, color = "#8B0000", fontface = "italic") +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12, color = "#8B0000"),
    plot.subtitle = element_text(hjust = 0.5, size = 8),
    legend.position = "right"
  )
```

## The Cost - Who Bears the Heaviest Burden?

#### Top 10 vs Bottom 10 States

```{r cost-who-bear, echo=FALSE, fig.width=11, fig.height=7}
#| message: false

df_div <- df %>%
  mutate(deviation = food_insecurity_proxy - nat_avg,
         direction = ifelse(deviation > 0, "Above Average", "Below Average")) %>%
  filter(rank(desc(deviation)) <= 10 | rank(deviation) <= 10) %>%
  mutate(state = fct_reorder(state, deviation))

ggplot(df_div, aes(x = deviation, y = state, fill = direction)) +
  geom_col(width = 0.5) +  
  geom_vline(xintercept = 0, color = "black", linewidth = 0.8) +
  geom_text(aes(label = sprintf("%+.1f", deviation), 
                hjust = ifelse(deviation > 0, -0.2, 1.2)),  # Adjusted position
            size = 3.5, fontface = "bold") +  # Increased size (was 2)
  scale_fill_manual(values = c("Above Average" = "#8B0000", "Below Average" = "#2980b9")) +
  labs(x = "Deviation from National Average (%)", y = NULL,
       title = "The Gap is a Policy Choice",
       subtitle = paste("National average:", round(nat_avg, 1), "% | Gap between worst and best:", 
                        round(max(df$food_insecurity_proxy) - min(df$food_insecurity_proxy), 1), "points")) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, color = "#8B0000", size = 13),
        legend.position = "top",
        axis.text.y = element_text(size = 9, face = "bold"))
```


## Intergenerational Impact - What Happens to Food Insecure Children?

#### 58% Require Government Support as Adults

```{r Insecure-children-cycle, echo=FALSE, fig.width=9, fig.height=5}
#| message: false

lifecycle <- data.frame(
  Stage = factor(c("Food Insecure\nChild", "Stunted\nDevelopment", "Lower\nEarnings", "Continued\nSupport"), 
                 levels = c("Food Insecure\nChild", "Stunted\nDevelopment", "Lower\nEarnings", "Continued\nSupport")),
  Probability = c(100, 68, 62, 58),
  Highlight = c(FALSE, FALSE, FALSE, TRUE)  # Only highlight final stage
)

ggplot(lifecycle, aes(x = Stage, y = Probability, fill = Highlight)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_text(aes(label = paste0(Probability, "%")), vjust = -0.5, size = 4, fontface = "bold") +
  scale_fill_manual(values = c("TRUE" = "#8B0000", "FALSE" = "#D3D3D3")) +
  labs(y = "Probability (%)", 
       title = "The Cost of Inaction",
       subtitle = "Without intervention, food insecure children become dependent adults") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "none", 
        plot.title = element_text(face = "bold", hjust = 0.5, color = "#8B0000", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, face = "bold"),
        axis.text.y = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())
```


## The Choice: Dependent Adults vs Self-Sufficient Citizens

```{r choice-comparison, echo=FALSE, fig.width=9, fig.height=5}
#| message: false

comparison <- data.frame(
  Outcome = c("Without Intervention", "With Early Intervention"),
  Percentage = c(58, 70)
)

ggplot(comparison, aes(x = Outcome, y = Percentage, fill = Outcome)) +
  geom_bar(stat = "identity", width = 0.4) +
  geom_text(aes(label = paste0(Percentage, "%")), vjust = -0.3, size = 4.5, fontface = "bold") +
  scale_fill_manual(values = c("Without Intervention" = "#8B0000", 
                                "With Early Intervention" = "#006400")) +
  labs(y = "Percentage of Children", 
       title = "Intervention Changes Everything",
       subtitle = "Without help: 58% dependent adults | With help: 70% self-sufficient citizens") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold", hjust = 0.5, color = "#8B0000", size = 12),
        plot.subtitle = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 10, face = "bold"))
```

## Unemployment = Hunger: States with high unemployment have double the food insecurity

```{r Unemployment-vs-Food-Insecurity, echo=FALSE, fig.width=8, fig.height=5}
#| message: false

ggplot(df, aes(x = unemployment_rate, y = food_insecurity_proxy)) +
  geom_point(color = "#8B0000", size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "#8B0000", linewidth = 1, fill = "#f0f0f0") +
  geom_text_repel(data = df %>% filter(unemployment_rate > 4.5 | food_insecurity_proxy > 11),
                  aes(label = state), size = 2.5, fontface = "bold") +
  labs(x = "Unemployment Rate (%)", 
       y = "Food Insecurity Rate (%)",
       title = "When Jobs Disappear, Hunger Arrives",
       subtitle = "States with >4.5% unemployment average 11%+ food insecurity") +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#8B0000", size = 12),
        plot.subtitle = element_text(hjust = 0.5, size = 8),
        axis.title.x = element_text(size = 9, margin = margin(t = 5)),
        axis.title.y = element_text(size = 9, margin = margin(r = 5)),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        plot.margin = margin(10, 10, 10, 10))
```

## Median Income vs Child Poverty — the inequality story

```{r Income-vs-ChildPoverty-simple, echo=FALSE, fig.width=10, fig.height=6}
#| message: false

ggplot(df, aes(x = median_income / 1000, y = poverty_children_u17)) +
  # Low income zone
  annotate("rect", xmin = -Inf, xmax = 65, ymin = 18, ymax = Inf,
           fill = "#c0392b", alpha = 0.08) +
  annotate("text", x = 55, y = 25, size = 3.5, hjust = 0,
           label = "Low income = High child poverty",
           color = "#c0392b", fontface = "bold") +
  # High income zone
  annotate("rect", xmin = 80, xmax = Inf, ymin = -Inf, ymax = 14,
           fill = "#27ae60", alpha = 0.08) +
  annotate("text", x = 81, y = 12, size = 3.5, hjust = 0,
           label = "High income = Low child poverty",
           color = "#27ae60", fontface = "bold") +
  # Trend line
  geom_smooth(method = "lm", se = TRUE, color = "#2c3e50",
              linetype = "dashed", linewidth = 1, fill = "grey85") +
  # Points - color only, no size
  geom_point(aes(color = food_insecurity_proxy), alpha = 0.85, size = 3) +
  # Label states
  geom_text_repel(
    data = df %>% filter(median_income < 65000 | poverty_children_u17 > 20),
    aes(label = abbr), size = 3, fontface = "bold",
    color = "#8B0000", box.padding = 0.4
  ) +
  scale_color_gradientn(name = "Food Insecurity\nRate (%)",
                        colors = c("#2ecc71", "#f39c12", "#c0392b")) +
  scale_x_continuous(labels = scales::dollar_format(suffix = "K")) +
  labs(
    title = "Wealth Protects — Poverty Sentences Children to Hunger",
    subtitle = "Lower income = higher child poverty = more food insecurity",
    x = "Median Household Income",
    y = "Child Poverty Rate (%)",
    caption = "Source: USDA SAIPE 2023"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 13, color = "#8B0000"),
    plot.subtitle = element_text(hjust = 0.5, size = 9, color = "#555555"),
    panel.grid.minor = element_blank()
  )
```


## Two Ways to See the Crisis: Rates vs Raw Numbers

#### Policy needs BOTH approaches: target high rates AND large populations

```{r two-perspectives-enhanced, echo=FALSE, fig.width=11, fig.height=6}
#| message: false

# Plot 1: Highest Poverty RATES (Hunger Belt)
p1 <- df %>%
  arrange(desc(poverty_children_u17)) %>%
  slice(1:6) %>%
  mutate(state = fct_reorder(state, poverty_children_u17)) %>%
  ggplot(aes(x = state, y = poverty_children_u17)) +
  geom_bar(stat = "identity", width = 0.6, fill = "#8B0000", alpha = 0.85) +
  geom_text(aes(label = paste0(round(poverty_children_u17, 1), "%")), 
            hjust = -0.05,  # Closer to bar (was -0.2)
            size = 2.8,     # Smaller text (was 4)
            fontface = "bold", 
            color = "#8B0000") +
  coord_flip(clip = "off") +
  labs(title = "HIGHEST CHILD POVERTY RATES", 
       subtitle = "The 'Hunger Belt' - Small states, severe crisis",
       x = NULL, 
       y = "Child Poverty Rate (%)") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 11, color = "#8B0000"),
        plot.subtitle = element_text(hjust = 0.5, size = 8, color = "#555555"),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10, face = "bold"),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.margin = margin(5, 15, 5, 5))

# Plot 2: Most Children in Poverty (RAW NUMBERS - Large States)
p2 <- df %>%
  arrange(desc(children_in_poverty)) %>%
  slice(1:6) %>%
  mutate(children_in_poverty_m = children_in_poverty / 1000000,
         state = fct_reorder(state, children_in_poverty_m)) %>%
  ggplot(aes(x = state, y = children_in_poverty_m)) +
  geom_bar(stat = "identity", width = 0.6, fill = "#1a5276", alpha = 0.85) +
  geom_text(aes(label = paste0(round(children_in_poverty_m, 1), "M")), 
            hjust = -0.05,  # Closer to bar
            size = 2.8,     # Smaller text
            fontface = "bold", 
            color = "#1a5276") +
  coord_flip(clip = "off") +
  labs(title = "MOST CHILDREN IN POVERTY", 
       subtitle = "Large states - Smaller rates, but millions affected",
       x = NULL, 
       y = "Children in Poverty (Millions)") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 11, color = "#1a5276"),
        plot.subtitle = element_text(hjust = 0.5, size = 8, color = "#555555"),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10, face = "bold"),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.margin = margin(5, 15, 5, 5))

# Combine plots
gridExtra::grid.arrange(p1, p2, ncol = 2)

# Add footer note
cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("KEY INSIGHT: Different problems need different solutions\n")
cat("• High RATE states (red) need targeted poverty programs\n")
cat("• High VOLUME states (blue) need scaled infrastructure & funding\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
```

## Key Findings: What The Data Tells Congress

#### Four Unavoidable Truths About Food Insecurity in America

```{r key-findings, echo=FALSE, fig.width=10, fig.height=6}
#| message: false

par(mfrow = c(2, 2), mar = c(3, 3, 3, 3), bg = "#f5f5f0", cex.main = 0.9)

# Finding 1: Correlation
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10))
rect(0.5, 0.5, 9.5, 9.5, col = "#FFFFFF", border = "#8B0000", lwd = 2)
text(5, 8, "FINDING 1", cex = 1.5, font = 2, col = "#8B0000")
text(5, 6.5, "Poverty = Hunger", cex = 1.2, font = 2)
text(5, 5, paste("r =", cor_val), cex = 2, font = 2, col = "#8B0000")
text(5, 3.5, "Near-perfect correlation", cex = 0.85)
text(5, 2.5, "Solve poverty = Solve hunger", cex = 0.85, font = 3)

# Finding 2: Hunger Belt
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10))
rect(0.5, 0.5, 9.5, 9.5, col = "#FFFFFF", border = "#8B0000", lwd = 2)
text(5, 8, "FINDING 2", cex = 1.5, font = 2, col = "#8B0000")
text(5, 6.5, "The 'Hunger Belt'", cex = 1.2, font = 2)
text(5, 5, "7 States", cex = 1.5, font = 2, col = "#8B0000")
text(5, 3.8, "LA, MS, AR, AL, KY, WV, OK", cex = 0.85)
text(5, 2.8, "3.4x higher food insecurity", cex = 0.85)
text(5, 1.8, "42% of food insecure children", cex = 0.85, col = "#8B0000")

# Finding 3: Intergenerational Cycle
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10))
rect(0.5, 0.5, 9.5, 9.5, col = "#FFFFFF", border = "#8B0000", lwd = 2)
text(5, 8, "FINDING 3", cex = 1.5, font = 2, col = "#8B0000")
text(5, 6.5, "The Pipeline to Dependency", cex = 1.2, font = 2)
text(5, 5, "58%", cex = 2, font = 2, col = "#8B0000")
text(5, 3.5, "require government support", cex = 0.85)
text(5, 2.5, "as adults without intervention", cex = 0.85)
text(5, 1.5, "70% become self-sufficient WITH help", cex = 0.85, font = 3, col = "#1a5276")

# Finding 4: Two Crises
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10))
rect(0.5, 0.5, 9.5, 9.5, col = "#FFFFFF", border = "#8B0000", lwd = 2)
text(5, 8, "FINDING 4", cex = 1.5, font = 2, col = "#8B0000")
text(5, 6.5, "Two Different Problems", cex = 1.2, font = 2)
text(5, 5, "High RATES", cex = 1.2, font = 2, col = "#8B0000")
text(5, 4.2, "Hunger Belt (7 states)", cex = 0.8)
text(5, 3.2, "High VOLUME", cex = 1.2, font = 2, col = "#1a5276")
text(5, 2.4, "CA, TX, FL, NY, IL, PA, OH", cex = 0.8)
text(5, 1.5, "5.2 million children in poverty", cex = 0.8, col = "#8B0000")
```

## Policy Recommendations

#### Three-Tier Response Based on Data

```{r policy, echo=FALSE}
#| message: false

recommendations <- data.frame(
  Priority = c("IMMEDIATE", "SHORT-TERM", "LONG-TERM"),
  Target = c("Hunger Belt (7 states)", "High-poverty districts", "High-volume states"),
  Action = c("Expand SNAP benefits\nEmergency food assistance", "Universal school breakfast\nSummer meal programs", "WIC expansion\nEarly childhood nutrition"),
  Expected_Impact = c("Reduce food insecurity 3-5%\nwithin 6 months", "Reach 5M+ children\n$3 return per $1 spent", "Break intergenerational cycle\nPrevent $3.8B annual costs")
)

knitr::kable(recommendations, 
             caption = "Evidence-Based Policy Response",
             format = "html") %>%
  kableExtra::kable_styling(font_size = 11, full_width = TRUE) %>%
  kableExtra::column_spec(1, bold = TRUE, color = c("#8B0000", "#e67e22", "#1a5276")) %>%
  kableExtra::row_spec(0, bold = TRUE, background = "#f0f0f0")
```

## 

```{r thank-you-slide, echo=FALSE, fig.width=12, fig.height=9}
#| message: false

par(mar = c(1, 1, 1, 1), bg = "#f5f5f0")
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10))

# Decorative line
lines(c(1, 9), c(9.2, 9.2), col = "#8B0000", lwd = 1)
lines(c(1, 9), c(9.0, 9.0), col = "#8B0000", lwd = 0.5)

# "THANK YOU" in center (INCREASED from 2.2 to 2.8)
text(5, 8.2, "THANK YOU", cex = 2.8, font = 2, col = "#8B0000")
text(5, 7.5, "Questions & Discussion", cex = 1.2, font = 3, col = "#555555")  # Increased from 1 to 1.2

# Decorative line below
lines(c(2, 8), c(7.0, 7.0), col = "#D3D3D3", lwd = 0.5)

# Data source (INCREASED from 0.7 to 0.9)
text(5, 6.2, "Data Source:", cex = 0.9, font = 2, col = "#333333")
text(5, 5.8, "USDA Economic Research Service & Census Bureau 2023", cex = 0.85, col = "#555555")  # Increased from 0.65

# Analysis (INCREASED from 0.7 to 0.9)
text(5, 5.2, "Analysis:", cex = 0.9, font = 2, col = "#333333")
text(5, 4.8, paste("51-state correlation analysis (r =", cor_val, ") between poverty and food insecurity"), 
     cex = 0.85, col = "#555555")  # Increased from 0.65

# Contact (INCREASED from 0.7 to 0.9)
text(5, 4.2, "Contact:", cex = 0.9, font = 2, col = "#333333")
text(5, 3.8, "Mehreen Ali Gillani | Policy Analysis Division", cex = 0.85, col = "#555555")  # Increased from 0.65

# Divider
lines(c(2.5, 7.5), c(3.3, 3.3), col = "#D3D3D3", lwd = 0.5, lty = 2)

# Closing quote (INCREASED from 0.65 to 0.85)
text(5, 2.7, "\"The children facing food insecurity today will become tomorrow's", 
     cex = 0.85, font = 3, col = "#8B0000")
text(5, 2.3, "dependent adults unless Congress acts now.", 
     cex = 0.85, font = 3, col = "#8B0000")
text(5, 1.9, "The data is clear. The solution is known. The only question is: will you act?\"", 
     cex = 0.85, font = 3, col = "#8B0000")

# Small red accent at bottom (removed alpha)
rect(4, 0.8, 6, 1.0, col = "#8B0000", border = NA)
```