---
title: "Story2_DATA608"
author: "Mehreen Ali Gillani"
date: "2026-02-21"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,           # Width in inches
  fig.height = 5,          # Height in inches
  fig.align = 'center',    # Center figures
  out.width = '100%',      # Use full text width
  dpi = 300) 
# Load required libraries
library(tidyverse)
library(plotly)
library(gridExtra)
library(RColorBrewer)
library(ragg)
```

# INTRODUCTION

The Federal Reserve operates under a dual mandate from Congress: **maintain price stability** (control inflation) and **promote maximum employment** (maintain low unemployment). These objectives often appear contradictory—policies that cool inflation (raising interest rates) can slow economic growth and increase unemployment, while policies that boost employment (lowering rates) can fuel inflation.​

*This assignment analyzes **25 years of economic data (1999-2024)** to answer a critical question: **"Has the Fed been able to fulfill the mandate given to it by Congress?"***

​

```{r}
# Read the data from Python
data <- read.csv("fed_mandate_complete_data.csv")
data$date <- as.Date(data$date)

# ============================================
# Ensure proper data types and recalculate goal columns
# ============================================

# Convert columns to proper numeric types
data$fed_funds_rate <- as.numeric(data$fed_funds_rate)
data$cpi <- as.numeric(data$cpi)
data$unemployment <- as.numeric(data$unemployment)
data$cpi_yoy <- as.numeric(data$cpi_yoy)

# Recalculate goal columns to ensure they're proper logical/numeric
data <- data %>%
  mutate(
    # Ensure these are numeric first
    cpi_yoy = as.numeric(cpi_yoy),
    unemployment = as.numeric(unemployment),
    
    # Recalculate goal columns
    inflation_goal_met = !is.na(cpi_yoy) & abs(cpi_yoy - 2) <= 1,
    unemployment_goal_met = !is.na(unemployment) & unemployment <= 5,
    both_goals_met = inflation_goal_met & unemployment_goal_met,
    
    # Convert to numeric for mean calculations (0/1)
    inflation_goal_num = as.numeric(inflation_goal_met),
    unemployment_goal_num = as.numeric(unemployment_goal_met),
    both_goals_num = as.numeric(both_goals_met),
    neither_goal_num = as.numeric(!inflation_goal_met & !unemployment_goal_met),
    
    year = as.numeric(year),
    decade = as.numeric(decade)
  )

# Remove any rows with NA values for key metrics
data_clean <- data %>% 
  filter(!is.na(cpi_yoy), !is.na(unemployment)) %>%
  mutate(
    decade = floor(year/10) * 10  # Recalculate decade to ensure consistency
  )

# ============================================
# VISUALIZATION 1: The Dual Mandate Timeline
# ============================================

# Create recession rectangles data frame
recessions <- data.frame(
  start = as.Date(c("2001-03-01", "2007-12-01", "2020-02-01")),
  end = as.Date(c("2001-11-01", "2009-06-01", "2020-04-01")),
  event = c("Dot-com Recession", "Great Recession", "COVID-19")
)

p1 <- ggplot(data_clean) +
  # Recession shading
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf, fill = event),
            alpha = 0.2) +
  # Main lines
  geom_line(aes(x = date, y = fed_funds_rate, color = "Federal Funds Rate"), size = 0.8) +
  geom_line(aes(x = date, y = unemployment, color = "Unemployment Rate"), size = 0.8) +
  geom_line(aes(x = date, y = cpi_yoy, color = "Inflation Rate"), size = 0.8) +
  # Target lines
  geom_hline(yintercept = 2, linetype = "dashed", color = "darkgreen", alpha = 0.7, size = 0.8) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "orange", alpha = 0.7, size = 0.8) +
  # Labels and theme
  labs(title = "The Fed's Dual Mandate: 25 Years of Balancing Inflation and Unemployment",
       subtitle = "Shaded areas indicate recessions | Dashed lines show target thresholds",
       x = "Year",
       y = "Rate (%)",
       color = "Indicator",
       fill = "Recession Period") +
  scale_color_manual(values = c("Federal Funds Rate" = "navy",
                                 "Unemployment Rate" = "darkgreen",
                                 "Inflation Rate" = "red")) +
  scale_fill_manual(values = c("Dot-com Recession" = "gray70",
                               "Great Recession" = "gray50",
                               "COVID-19" = "gray30")) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "gray40"),
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  ylim(-2, 15)

# Display
print(p1)
ggsave("visualization_1_dual_mandate_timeline.png", p1, width = 16, height = 10, dpi = 300)



# ============================================
# VISUALIZATION 2: Dot Plot with Smart Label Positioning
# ============================================

# Calculate performance by decade
performance_decade <- data_clean %>%
  filter(year >= 2000) %>%
  group_by(decade) %>%
  summarise(
    inflation_met = mean(inflation_goal_num, na.rm = TRUE) * 100,
    unemployment_met = mean(unemployment_goal_num, na.rm = TRUE) * 100,
    both_met = mean(both_goals_num, na.rm = TRUE) * 100,
    .groups = "drop"
  )

# Create a longer format for easier plotting
performance_long <- performance_decade %>%
  pivot_longer(cols = c(inflation_met, unemployment_met, both_met),
               names_to = "metric", values_to = "percentage") %>%
  mutate(
    metric_label = case_when(
      metric == "inflation_met" ~ "Inflation Goal",
      metric == "unemployment_met" ~ "Unemployment Goal",
      metric == "both_met" ~ "Both Goals"
    ),
    # Define label positions: left for unemployment, right for both, center for inflation
    label_hjust = case_when(
      metric == "unemployment_met" ~ 1.2,  # Left side
      metric == "both_met" ~ -0.2,          # Right side
      TRUE ~ 0.5                             # Center for inflation
    ),
    # Adjust vertical position slightly to avoid overlap
    label_vjust = case_when(
      decade == 2000 & metric == "inflation_met" ~ -1.2,
      decade == 2000 & metric == "unemployment_met" ~ 1.2,
      decade == 2000 & metric == "both_met" ~ -0.8,
      decade == 2010 & metric == "inflation_met" ~ -1.2,
      decade == 2010 & metric == "unemployment_met" ~ 1.2,
      decade == 2010 & metric == "both_met" ~ -1.2,
      decade == 2020 & metric == "inflation_met" ~ -1.2,
      decade == 2020 & metric == "unemployment_met" ~ -1.2,
      decade == 2020 & metric == "both_met" ~ 1.2,
      TRUE ~ -0.8
    )
  )

# Create dot plot with custom colors and smart label positioning
p2_alt <- ggplot() +
  # Add connecting lines (using the wide format for simplicity)
  geom_segment(data = performance_decade,
               aes(x = factor(decade), xend = factor(decade),
                   y = 0, yend = inflation_met), 
               color = "#E41A1C", alpha = 0.2, size = 1) +  # Softer red
  geom_segment(data = performance_decade,
               aes(x = factor(decade), xend = factor(decade),
                   y = 0, yend = unemployment_met), 
               color = "#4DAF4A", alpha = 0.2, size = 1) +  # Softer green
  geom_segment(data = performance_decade,
               aes(x = factor(decade), xend = factor(decade),
                   y = 0, yend = both_met), 
               color = "#984EA3", alpha = 0.2, size = 1) +  # Softer purple
  
  # Add points with better colors
  geom_point(data = filter(performance_long, metric == "inflation_met"),
             aes(x = factor(decade), y = percentage, color = "Inflation Goal"), 
             size = 5) +
  geom_point(data = filter(performance_long, metric == "unemployment_met"),
             aes(x = factor(decade), y = percentage, color = "Unemployment Goal"), 
             size = 5) +
  geom_point(data = filter(performance_long, metric == "both_met"),
             aes(x = factor(decade), y = percentage, color = "Both Goals"), 
             size = 5) +
  
  # Add labels with custom positioning
  geom_text(data = performance_long,
            aes(x = factor(decade), y = percentage, 
                label = sprintf("%.1f%%", percentage),
                hjust = label_hjust, vjust = label_vjust,
                color = metric_label),
            size = 3.5, show.legend = FALSE) +
  
  # Labels and theme
  labs(title = "Fed Mandate Performance by Decade",
       subtitle = "Height shows % of months meeting each goal",
       x = "Decade",
       y = "Percentage of Months Meeting Goal (%)",
       color = "Performance Metric") +
  
  # Custom color palette
  scale_color_manual(values = c("Inflation Goal" = "darkred",      # Bright red
                                 "Unemployment Goal" = "darkgreen",  # Bright green
                                 "Both Goals" = "#984EA3")) +      # Bright purple
  
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 13, color = "gray40"),
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    legend.position = "bottom",
    legend.text = element_text(size = 13),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray90", linetype = "dotted"),
    plot.margin = margin(10, 30, 10, 30)  # Extra margin for labels
  ) +
  ylim(0, 100) +
  scale_x_discrete(expand = expansion(mult = 0.15))  # More padding for labels

print(p2_alt)
ggsave("visualization_2_dot_plot_clean.png", p2_alt, width = 16, height = 10,units = "in", dpi = 300)

# ============================================
# VISUALIZATION 3: The Trade-off Scatter Plot
# ============================================

p3 <- ggplot(data_clean %>% filter(!is.na(cpi_yoy), !is.na(unemployment)), 
             aes(x = unemployment, y = cpi_yoy, color = factor(decade))) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 0.5) +
  labs(title = "The Inflation-Unemployment Trade-off (Phillips Curve)",
       subtitle = "Each point represents one month (1999-2024)",
       x = "Unemployment Rate (%)",
       y = "Inflation Rate (CPI YoY %)",
       color = "Decade") +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "bottom"
  ) +
  geom_hline(yintercept = 2, linetype = "dotted", color = "darkgreen", alpha = 0.5) +
  geom_vline(xintercept = 5, linetype = "dotted", color = "orange", alpha = 0.5) +
  annotate("text", x = 3, y = 8, label = "Target Zone", color = "darkgreen", size = 3)

print(p3)
ggsave("visualization_3_phillips_curve.png", p3, width = 16, height = 10, dpi = 300)

# ============================================
# SUMMARY STATISTICS FOR YOUR STORY 
# ============================================

cat("\n", rep("=", 60), "\n", sep="")
cat("FED MANDATE PERFORMANCE SUMMARY (1999-2024)")
cat("\n", rep("=", 60), "\n", sep="")

# Use the numeric columns for calculations
overall <- data_clean %>%
  summarise(
    `Total Months` = n(),
    `Avg Inflation` = sprintf("%.2f%%", mean(cpi_yoy, na.rm = TRUE)),
    `Avg Unemployment` = sprintf("%.2f%%", mean(unemployment, na.rm = TRUE)),
    `Avg Fed Funds Rate` = sprintf("%.2f%%", mean(fed_funds_rate, na.rm = TRUE)),
    `Inflation Goal Met` = sprintf("%.1f%%", mean(inflation_goal_num, na.rm = TRUE) * 100),
    `Unemployment Goal Met` = sprintf("%.1f%%", mean(unemployment_goal_num, na.rm = TRUE) * 100),
    `Both Goals Met` = sprintf("%.1f%%", mean(both_goals_num, na.rm = TRUE) * 100),
    `Neither Goal Met` = sprintf("%.1f%%", mean(neither_goal_num, na.rm = TRUE) * 100)
  )

print(overall)

cat("\n", rep("=", 60), "\n", sep="")
cat("DECADE-BY-DECADE BREAKDOWN")
cat("\n", rep("=", 60), "\n", sep="")

# Use numeric columns for decade summary
decade_summary <- data_clean %>%
  filter(decade >= 2000) %>%
  group_by(decade) %>%
  summarise(
    `Avg Inflation` = sprintf("%.2f%%", mean(cpi_yoy, na.rm = TRUE)),
    `Avg Unemployment` = sprintf("%.2f%%", mean(unemployment, na.rm = TRUE)),
    `Inflation Goal` = sprintf("%.1f%%", mean(inflation_goal_num, na.rm = TRUE) * 100),
    `Unemployment Goal` = sprintf("%.1f%%", mean(unemployment_goal_num, na.rm = TRUE) * 100),
    `Both Goals` = sprintf("%.1f%%", mean(both_goals_num, na.rm = TRUE) * 100),
    .groups = "drop"
  )

print(decade_summary)

cat("\n All visualizations created successfully!\n")
```

```{r}
# ============================================

# VISUALIZATION 4: Lagged Effects Analysis
# ============================================

# Calculate cross-correlations at different lags
max_lag <- 24  # Look at up to 24 months of lags

# Function to calculate correlation at different lags
lag_correlation <- function(x, y, max_lag) {
  result <- data.frame(
    lag = -max_lag:max_lag,
    correlation = NA
  )
  
  for (i in -max_lag:max_lag) {
    if (i < 0) {
      # Negative lag: y leads x (x lags behind y)
      y_lead <- lead(y, abs(i))
      result[result$lag == i, "correlation"] <- cor(x, y_lead, use = "complete.obs")
    } else if (i > 0) {
      # Positive lag: x leads y (y lags behind x)
      x_lead <- lead(x, i)
      result[result$lag == i, "correlation"] <- cor(x_lead, y, use = "complete.obs")
    } else {
      # Zero lag: contemporaneous correlation
      result[result$lag == 0, "correlation"] <- cor(x, y, use = "complete.obs")
    }
  }
  return(result)
}

# Calculate lag correlations
fed_inflation_lag <- lag_correlation(data_clean$fed_funds_rate, 
                                      data_clean$cpi_yoy, max_lag)
fed_unemployment_lag <- lag_correlation(data_clean$fed_funds_rate, 
                                         data_clean$unemployment, max_lag)

# Add relationship type
fed_inflation_lag$relationship <- "Fed Rate vs Inflation"
fed_unemployment_lag$relationship <- "Fed Rate vs Unemployment"

# Combine
lag_data <- rbind(fed_inflation_lag, fed_unemployment_lag)

# Find peak lags
peak_lags <- lag_data %>%
  group_by(relationship) %>%
  slice_max(abs(correlation), n = 1) %>%
  mutate(
    direction = ifelse(lag > 0, "Fed leads", ifelse(lag < 0, "Fed lags", "Simultaneous")),
    lag_months = abs(lag),
    lag_text = paste0("Peak at lag ", lag_months, " months\n(Fed ", direction, ")")
  )

# Create heatmap
p4 <- ggplot(lag_data, aes(x = lag, y = relationship, fill = correlation)) +
  geom_tile() +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", 
                       midpoint = 0, limit = c(-1, 1),
                       name = "Correlation") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", size = 0.5) +
  geom_text(data = peak_lags, 
            aes(x = lag, y = relationship, label = sprintf("%.2f", correlation)),
            color = "black", fontface = "bold", size = 4) +
  labs(title = "When Do Fed Actions Show Their Effect?",
       subtitle = "Cross-correlation at different time lags | Positive lag = Fed rate change comes first",
       x = "Lag (months) - Positive means Fed rate leads",
       y = "",
       caption = "Interpretation: Positive lag means Fed rate change precedes the effect") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "gray40"),
    axis.text.y = element_text(size = 11),
    legend.position = "bottom",
    legend.key.width = unit(2, "cm")
  ) +
  scale_x_continuous(breaks = seq(-24, 24, 6))

print(p4)
#ggsave("visualization_4_lag_analysis.png", p4, width = 16, height = 9, dpi = 300)

```

```{r}
# ============================================
# VISUALIZATION 5: Rate Hike Episodes Analysis
# ============================================

# Load lubridate explicitly
library(lubridate)

# Select major hiking episodes
hiking_episodes <- data.frame(
  episode = c("2004-2006 Tightening", "2015-2018 Normalization", "2022-2023 Inflation Fight"),
  start_date = as.Date(c("2004-06-01", "2015-12-01", "2022-03-01")),
  end_date = as.Date(c("2006-06-01", "2018-12-01", "2023-07-01"))
)

# Create event study data
event_study <- data.frame()

for(i in 1:nrow(hiking_episodes)) {
  episode <- hiking_episodes[i, ]
  cat("Processing:", episode$episode, "\n")
  
  # Filter data for this episode
  episode_data <- data_clean %>%
    filter(date >= episode$start_date - 365,  # 12 months before
           date <= episode$start_date + 730)   # 24 months after
  
  if(nrow(episode_data) > 0) {
    episode_data <- episode_data %>%
      mutate(
        # Calculate months since start
        months_since_start = as.numeric(difftime(date, episode$start_date, units = "days")) / 30.44,
        episode = episode$episode
      )
    
    event_study <- bind_rows(event_study, episode_data)
  }
}

# Check if we have data
if(nrow(event_study) > 0) {
  # Normalize values to start of episode
  event_study_normalized <- event_study %>%
    group_by(episode) %>%
    mutate(
      # Find the index at month 0 (closest to start)
      start_idx = which.min(abs(months_since_start)),
      fed_start = fed_funds_rate[start_idx],
      cpi_start = cpi[start_idx],
      unemp_start = unemployment[start_idx],
      
      # Create normalized values
      fed_norm = (fed_funds_rate / fed_start) * 100,
      cpi_norm = (cpi / cpi_start) * 100,
      unemp_norm = (unemployment / unemp_start) * 100
    ) %>%
    filter(months_since_start >= -11, months_since_start <= 23)  # Clean range
  
  # Create the plot
  p5 <- ggplot(event_study_normalized, aes(x = months_since_start, color = episode)) +
    geom_line(aes(y = fed_norm, linetype = "Fed Funds Rate"), size = 1) +
    geom_line(aes(y = cpi_norm, linetype = "CPI (Inflation)"), size = 1) +
    geom_line(aes(y = unemp_norm, linetype = "Unemployment"), size = 1) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "black", alpha = 0.5) +
    labs(title = "What Happens After the Fed Starts Raising Rates?",
         subtitle = "Indexed to 100 at start of hiking cycle",
         x = "Months Since First Rate Hike",
         y = "Index (Start of Hike = 100)",
         color = "Hiking Episode",
         linetype = "Indicator") +
    scale_x_continuous(breaks = seq(-12, 24, 6)) +
    scale_color_brewer(palette = "Set1") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 10, color = "gray40"),
      legend.position = "bottom",
      legend.box = "vertical"
    ) +
    geom_hline(yintercept = 100, linetype = "dotted", alpha = 0.3)
  
  print(p5)
  ggsave("visualization_5_hiking_episodes.png", p5, width = 18, height = 10, dpi = 300)
  
} else {
  print("Not enough data for episode analysis")
}

```

### The Ripple Effect: How Fed Policy Propagates Through the Economy

```{r lag-analysis, echo=FALSE, message=FALSE, warning=FALSE}
# Create lagged versions of variables
lead_lag_data <- data_clean %>%
  select(date, fed_funds_rate, cpi_yoy, unemployment) %>%
  mutate(
    # Fed leads (Fed rate now vs. future inflation/unemployment)
    fed_lead_6m_inflation = lead(cpi_yoy, 6),
    fed_lead_12m_inflation = lead(cpi_yoy, 12),
    fed_lead_6m_unemp = lead(unemployment, 6),
    fed_lead_12m_unemp = lead(unemployment, 12),
    
    # Fed lags (past inflation/unemployment vs. current Fed rate)
    inflation_lag_6m = lag(cpi_yoy, 6),
    inflation_lag_12m = lag(cpi_yoy, 12),
    unemp_lag_6m = lag(unemployment, 6),
    unemp_lag_12m = lag(unemployment, 12)
  )

# Create scatter plot matrix
library(GGally)

# Select relevant columns for matrix
matrix_data <- lead_lag_data %>%
  select(
    `Fed Rate` = fed_funds_rate,
    `Inflation (current)` = cpi_yoy,
    `Unemployment (current)` = unemployment,
    `Inflation in 12mo` = fed_lead_12m_inflation,
    `Unemployment in 12mo` = fed_lead_12m_unemp
  ) %>%
  na.omit()

# Custom colors for the plot
custom_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00")

# Create the plot with updated colors and sizes
p6 <- ggpairs(matrix_data,
              title = "Lead-Lag Relationships: Fed Actions and Economic Outcomes",
              
              # Upper triangle (correlations)
              upper = list(
                continuous = wrap("cor", 
                                  size = 6,           # Larger correlation text
                                  color = "black",    # Black text
                                  fontface = "bold")  # Bold text
              ),
              
              # Lower triangle (scatter plots)
              lower = list(
                continuous = wrap("points", 
                                  alpha = 0.6,        # Slightly transparent
                                  size = 1.5,         # Larger points
                                  color = "#1f77b4")  # Nice blue color
              ),
              
              # Diagonal (density plots)
              diag = list(
                continuous = wrap("barDiag", 
                                  bins = 30,
                                  fill = "#E41A1C",   # Red fill
                                  color = "black",    # Black outline
                                  alpha = 0.7)        # Slight transparency
              )
) +
  
theme_minimal() 

# Save for PPT with high resolution
ggsave("ppt_lead_lag_matrix_large.png", 
       plot = p6, 
       width = 16, 
       height = 14, 
       units = "in", 
       dpi = 300, 
       bg = "white")

print(p6)
 
```

### The Fed's casual chain

```{r}
# ============================================
# VISUALIZATION 7: Annotated Timeline Showing Causal Chain 
# ============================================

# Extend data through 2008 to show the full cycle
p7_data <- data_clean %>% 
  filter(date >= as.Date("2003-09-01"), year(date) <= 2008)  # Extended to 2008

# Find the exact peak values for accurate label placement
peak_inflation_date <- p7_data$date[which.max(p7_data$cpi_yoy)]
peak_inflation_value <- max(p7_data$cpi_yoy, na.rm = TRUE)
peak_unemp_date <- p7_data$date[which.max(p7_data$unemployment)]
peak_unemp_value <- max(p7_data$unemployment, na.rm = TRUE)

# Create annotations dataframe for cleaner code
annotations <- data.frame(
  label = c(
    "Fed raises rates\n(2004-2006)",
    "Inflation peaks\n~15 months later",
    "Unemployment rises\n~21 months later",
    "Financial Crisis\nbegins"
  ),
  date = as.Date(c(
    "2004-12-01",
    "2005-09-01", 
    "2006-09-01", 
    "2008-09-15"
  )),
  y = c(6.2, 2.8, 6.8, 8.5),  # Repositioned: inflation label lower (2.8)
  color = c("navy", "red", "darkgreen", "gray40"),
  hjust = c(0.3, 0.3, 0.3, 0.5)
)

p7 <- ggplot(p7_data) +
  # Main lines
  geom_line(aes(x = date, y = fed_funds_rate, color = "Federal Funds Rate"), size = 1.2) +
  geom_line(aes(x = date, y = cpi_yoy, color = "Inflation Rate"), size = 1.2) +
  geom_line(aes(x = date, y = unemployment, color = "Unemployment Rate"), size = 1.2) +
  
  # Annotate the lags with arrows
  
  # Add vertical line for crisis
  geom_vline(xintercept = as.Date("2008-09-15"), 
             linetype = "dotted", color = "gray40", size = 0.8) +
  
  # Add labels with positioning
  geom_label(data = annotations[1,],  # Fed rate label
            aes(x = date, y = y, label = label),
            fill = "navy", color = "white", alpha = 0.8,
            size = 3.0, fontface = "bold", hjust = 0.5) +
  
  geom_label(data = annotations[2,],  # Inflation label (now LOWER)
            aes(x = date, y = y, label = label),
            fill = "red", color = "white", alpha = 0.8,
            size = 3.0, fontface = "bold", hjust = 0.5) +
  
  geom_label(data = annotations[3,],  # Unemployment label
            aes(x = date, y = y, label = label),
            fill = "darkgreen", color = "white", alpha = 0.8,
            size = 3.0, fontface = "bold", hjust = 0.5) +
  
  geom_label(data = annotations[4,],  # Crisis label
            aes(x = date, y = y, label = label),
            fill = "gray40", color = "white", alpha = 0.8,
            size = 3.0, fontface = "bold", hjust = 0.5) +
  
  # Add subtle points at peaks for clarity
  geom_point(data = p7_data %>% filter(date == peak_inflation_date),
             aes(x = date, y = cpi_yoy), color = "red", size = 3) +
  geom_point(data = p7_data %>% filter(date == peak_unemp_date),
             aes(x = date, y = unemployment), color = "darkgreen", size = 3) +
  
  # Labels and theme
  labs(title = "The Fed's Causal Chain: Rate Hikes → Lower Inflation → Higher Unemployment",
       subtitle = "2004-2008 cycle shows typical time lags | Inflation peaks before unemployment rises",
       x = "Year",
       y = "Rate (%)",
       color = "Indicator",
       caption = "Gray arrows show causal sequence | Dotted line marks Financial Crisis") +
  
  scale_color_manual(values = c("Federal Funds Rate" = "navy",
                                 "Inflation Rate" = "red",
                                 "Unemployment Rate" = "darkgreen")) +
  
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  
theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

print(p7)

```

### Conclusion:

The Fed has been a capable crisis manager but an inconsistent mandate fulfiller. The dual mandate remains an aspirational target—achievable only about a quarter of the time—rather than a consistently attainable goal.​

#### **Recommendations for Policymakers**​

-   Communicate the lag effect more clearly to the public​

-   Set realistic expectations that both goals cannot always be met​

-   Consider a flexible timeline for achieving both mandates rather than simultaneous success​

*"Monetary policy is a balancing act—and the Fed walks a tightrope between inflation and unemployment every single day."*

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
